Conditional selectivity for triplestores

### 2012-09-18 conditional/trunk/selectivity-basic.tex
cutnpasto
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@375 43ef1190-90fd-4dd7-a769-d0121f02b80e
cutnpasto
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@376 43ef1190-90fd-4dd7-a769-d0121f02b80e
void:triples -> void:entities
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@377 43ef1190-90fd-4dd7-a769-d0121f02b80e

We note that VoID includes the concept of \emph{property partitions}
(see section~4.5 in \cite{voidnote}): ``A property-based partition
contains only those triples of a dataset that use a particular
predicate.'', and since a partition is itself a dataset, the same
statistical properties can be applied, such as \rdfterm{void:triples},
\rdfterm{void:distinctSubjects} and \rdfterm{void:distinctObjects}.

In the case where the predicate is known, we have
\begin{equation}\label{eq:selPp}
 \sel{P = p} =
   \frac{\rdfterm{void:propertyPartition/void:triples}}{|d|} ,
\end{equation}
where the notation is inspired by SPARQL 1.1 Property Paths, for
example the relevant part of a VoID description for the second example
in section~\ref{sec:example} would be:
\begin{verbatim}
 </dataset> a void:Dataset ;
            void:propertyPartition [
                void:property foaf:firstName;
                void:triples 2 .
            ]
\end{verbatim}
where $p = \rdfterm{foaf:firstName}$.

In the case where there is a given predicate but the subject is a
variable, we can use the \rdfterm{void:distinctSubjects} property:
\begin{equation}\label{eq:selSvPp}
 \sel{S = ? | P = p} =
   \frac{\rdfterm{void:propertyPartition/void:distinctSubjects}}{\rdfterm{void:propertyPartition/void:triples}} ,
\end{equation}
and similarly where the object is a variable:
\begin{equation}\label{eq:selOvPp}
 \sel{O = ? | P = p} =
   \frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{\rdfterm{void:propertyPartition/void:triples}} .
\end{equation}

We also have in the special case of a given class, i.e. predicate
\rdfterm{rdf:type} or $P = a$ and $O = c$, we can use
\rdfterm{void:classPartition} (also defined in section~4.5 in
\cite{voidnote}):
\begin{equation}\label{eq:selSvPaOc}
 \sel{S = ? | P = a \cap O = c } =
    \frac{\rdfterm{void:classPartition/void:triples}}{|d|} .
\end{equation}

@@

This case is similar to the above, and we can write 
\begin{equation}\label{eq:selsuobunpb}
\sel{T_{\{?~p~?\}}} = \sel{O = ? | P = p \cap S = s} \cdot 
\sel{S = s | P = p} \cdot
\sel{P = p} . 
\end{equation}
We note that we can't find the first two parts in the VoID
description, but we also note with interest that the first part shows
that only the marginal distribution of objects is needed to estimate
this part.

>>>>

We note that VoID includes the concept of \emph{property partitions}
(see section~4.5 in \cite{voidnote}): ``A property-based partition
contains only those triples of a dataset that use a particular
predicate.'', and since a partition is itself a dataset, the same
statistical properties can be applied, such as \rdfterm{void:entities},
\rdfterm{void:distinctSubjects} and \rdfterm{void:distinctObjects}.

In the case where the predicate is known, we have
\begin{equation}\label{eq:selPp}
 \sel{P = p} =
   \frac{\rdfterm{void:propertyPartition/void:entities}}{|d|} ,
\end{equation}
where the notation is inspired by SPARQL 1.1 Property Paths, for
example the relevant part of a VoID description for the second example
in section~\ref{sec:example} would be:
\begin{verbatim}
 </dataset> a void:Dataset ;
            void:propertyPartition [
                void:property foaf:firstName;
                void:entities 2 .
            ]
\end{verbatim}
where $p = \rdfterm{foaf:firstName}$.

In the case where there is a given predicate but the subject is a
variable, we can use the \rdfterm{void:distinctSubjects} property:
\begin{equation}\label{eq:selSvPp}
 \sel{S = ? | P = p} =
   \frac{\rdfterm{void:propertyPartition/void:distinctSubjects}}{\rdfterm{void:propertyPartition/void:entities}} ,
\end{equation}
and similarly where the object is a variable:
\begin{equation}\label{eq:selOvPp}
 \sel{O = ? | P = p} =
   \frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{\rdfterm{void:propertyPartition/void:entities}} .
\end{equation}

We also have in the special case of a given class, i.e. predicate
\rdfterm{rdf:type} or $P = a$ and $O = c$, we can use
\rdfterm{void:classPartition} (also defined in section~4.5 in
\cite{voidnote}):
\begin{equation}\label{eq:selSvPaOc}
 \sel{S = ? | P = a \cap O = c } =
    \frac{\rdfterm{void:classPartition/void:entities}}{|d|} .
\end{equation}

@@

This case is similar to the above, and we can write 
\begin{equation}\label{eq:selsuunobpb}
\sel{T_{\{s~p~?\}}} = \sel{O = ? | P = p \cap S = s} \cdot 
\sel{S = s | P = p} \cdot
\sel{P = p} . 
\end{equation}
We note that we can't find the first two parts in the VoID
description, but we also note with interest that the first part shows
that only the marginal distribution of objects is needed to estimate
this part.
